Lab 4: Graphical interpretation of LMs: Answer Key

### 2023-01-17 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM.Rmd
Add files via upload

#Installing and loading packages for use
pacman::p_load(tidyverse,palmerpenguins,ggeffects)

@@

```{r}
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(size = cyl), colour = "red") +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_bw()

ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_bw()
```

@@

@@

## Interaction analysis
+ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids.

>>>>

#Installing and loading packages for use
pacman::p_load(tidyverse,palmerpenguins,ggeffects,ggthemes)

@@

```{r}
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(size = cyl), colour = "red") +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders")

ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") 
```

@@

## Brief divergence 1:Themes

+ As you've hopefully seen ggplot is highly customization. One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g. color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code. This is what we were doing using theme_bw() above.

+ You can also create custom themes yourself!

```{r}
#Here is a theme I use often

plot_theme = theme(text = element_text(size=11),  #set text size to 11
                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
                   panel.grid.major = element_blank(),
                   panel.background = element_blank(),#these lines make the background of the plot blank
                   plot.background = element_rect(fill = "transparent",colour = NA),
                   axis.text = element_text(color="black"), #setting axis text to black 
                   axis.line = element_line(colour = "black"),#setting axis lines to black 
                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
)

#Using the theme
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  plot_theme

#As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
#bw
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_bw()

#minimal
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_minimal()

#classic
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_classic()

#If you want even more themes you can use the ggthemes library!
#install.packages("ggthemes")
library(ggthemes)

#WSJ
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_wsj()

#538
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_fivethirtyeight()

#The Economist
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_economist()

@@

## Brief divergence 2: Facets

+ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.

1. facet_wrap() "wraps" a 1d ribbon of panels into 2d. Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data). This is most useful when you want to facet by a single variable (or two)

carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_wrap(~ am)
```

2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.

```{r}
#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.

carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(vs ~ am)

carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(am ~ .)

carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(. ~ am)
```

## Interaction analysis
+ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids (like facets).

### 2023-01-17 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM_Answers.Rmd
Add files via upload

@@

#Visualize the results in a coefficient plot
Penguin_results <- ggplot(data=temp_lm, aes(x=term, y=estimate)) + 
    geom_hline(yintercept=0, color="red", size=.5) +
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), size=.5, position=position_dodge(width=0.5)) +
    geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
    ylab("Effect of Flipper Length on Penguin Body Mass") +
    xlab("")+
    coord_flip()+
  geom_text(aes(y=estimate,label=round(estimate, digits=2)), size=2.5, vjust=-1.5, position = position_dodge(width=.2))

@@

## Interaction analysis
+ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids.

@@

#Plotting (make sure to assign color and shape to the model variable in the ggplot call)
sleep_results_plot <- ggplot(data=sleep_results, aes(x=term, y=estimate, 
           color = model, shape = model)) + 
    geom_hline(yintercept=0, color="red", size=.5) +
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), linewidth=.5, position=position_dodge(width=0.5)) +
    geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
  labs(title = "Model Estimates of Brain and Body Weight (and total sleep) on REM Sleep",
       x = "Predictor",
       y = "Coefficient Estimate",
       caption = "Models fit with OLS. Error bars show the 95% confidence interval.",
       color = "Model:",
       shape = "Model:") +
   scale_x_discrete(labels = c("Body Weight", "Brain Weight", "Total Sleep"))+ 
  coord_flip()+
  plot_theme

>>>>

## Themes

+ As you've hopefully seen ggplot is highly customization. One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g. color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code. This is what we were doing using theme_bw() above.

+ You can also create custom themes yourself!

```{r}
#Here is a theme I use often

plot_theme = theme(text = element_text(size=11),  #set text size to 11
                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
                   panel.grid.major = element_blank(),
                   panel.background = element_blank(),#these lines make the background of the plot blank
                   plot.background = element_rect(fill = "transparent",colour = NA),
                   axis.text = element_text(color="black"), #setting axis text to black 
                   axis.line = element_line(colour = "black"),#setting axis lines to black 
                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
)

#Using the theme
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  plot_theme

#As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
#bw
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_bw()

#minimal
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_minimal()

#classic
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_classic()

#If you want even more themes you can use the ggthemes library!
#install.packages("ggthemes")
library(ggthemes)

#WSJ
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_wsj()

#538
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_fivethirtyeight()

#The Economist
ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_economist()

@@

#Visualize the results in a coefficient plot
Penguin_results <- ggplot(data=temp_lm, aes(x=term, y=estimate)) + 
    geom_hline(yintercept=0, color="red", linewidth=.5) +
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), size=.5, position=position_dodge(width=0.5)) +
    geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
    ylab("Effect of Flipper Length on Penguin Body Mass") +
    xlab("")+
    coord_flip()+
  geom_text(aes(y=estimate,label=round(estimate, digits=2)), size=2.5, vjust=-1.5, position = position_dodge(width=.2))

@@

## Facets

+ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.

1. facet_wrap() "wraps" a 1d ribbon of panels into 2d. Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data). This is most useful when you want to facet by a single variable (or two)

carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_wrap(~ am)
```

2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.

```{r}
#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.

carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(vs ~ am)

carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(am ~ .)

carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(. ~ am)
```

## Interaction analysis
+ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids (like facets).

@@

#Plotting (make sure to assign color and shape to the model variable in the ggplot call)
sleep_results_plot <- ggplot(data=sleep_results, aes(x=term, y=estimate, 
           color = model, shape = model)) + 
    geom_hline(yintercept=0, color="red", size=.5) +
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), linewidth=.5, position=position_dodge(width=0.5)) +
    geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
  labs(title = "Model Estimates of Brain and Body Weight (and total sleep) on REM Sleep",
       x = "Predictor",
       y = "Coefficient Estimate",
       caption = "Models fit with OLS. Error bars show the 95% confidence interval.",
       color = "Model:",
       shape = "Model:") +
   scale_x_discrete(labels = c("Body Weight", "Brain Weight", "Total Sleep"))+ 
  coord_flip()+
  geom_text(aes(y=estimate,label=round(estimate, digits=2)), size=2.5, vjust=-1, position = position_dodge(width=.2))+
  plot_theme
